---
title: "R Notebook for modelling figures"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: console
---

# Set working directory
```{r}
# setwd("TYPE_PATH_TO/r_notebook_modelling/")
```

# Load libraries

```{r}

library(ggplot2)
library(ggpubr)
library(dplyr)
library(tidyr)

```

# colours for figures

```{r}

color_at_centre = "#E8CAAD"
color_at_margin = "#C45A11"

color_less_necrotic = "#C1FFC1"
color_more_necrotic = "#698B69"

```


# Figure panel (f) -- compare number of SCNAs at the centre vs at the margin

# input data : ./data_frames/df_necrotic_frac_scna_burden_all_v2.csv

```{r}
# --- read data frame ----
fin_name <- "./data_frames/df_necrotic_frac_scna_burden_all_v2.csv"
df_necrotic_frac_scna_burden_all_v2 <- read.csv(fin_name)

# --- filter out too necrotic regions & split into "less necrotic" vs "more necrotic" ----
df_modified <- df_necrotic_frac_scna_burden_all_v2[
  which(df_necrotic_frac_scna_burden_all_v2$Necrotic.Fraction < 0.8), 
]
median_necrotic_fraction <- median(df_modified$Necrotic.Fraction)
print(median_necrotic_fraction)

df_modified <- df_modified %>% mutate(
  Necrotic.Fraction.Interval = ifelse(
    df_modified$Necrotic.Fraction <= median_necrotic_fraction,
    "Less necrotic", "More necrotic"
  )
)

theme_set(theme_classic())
g <- ggplot(df_modified, aes(Location, Num.Driver.SCNA)) 

# --- print median & number of regions ----
print(median(df_modified[which(df_modified$Location == "At.Centre"), "Num.Driver.SCNA"]))
print(median(df_modified[which(df_modified$Location == "At.Margin"), "Num.Driver.SCNA"]))
print(mean(df_modified[which(df_modified$Location == "At.Centre"), "Num.Driver.SCNA"]))
print(mean(df_modified[which(df_modified$Location == "At.Margin"), "Num.Driver.SCNA"]))
print(nrow(df_modified[which(df_modified$Location == "At.Centre"), ]))
print(nrow(df_modified[which(df_modified$Location == "At.Margin"), ]))

# --- VIOLIN ----
p <- g +
  geom_violin(aes(fill= Location), draw_quantiles = c(0.25, 0.5, 0.75)) +
  #stat_summary(fun.y=mean, fill='black', color='black',geom='point', size=5) +
  #geom_jitter(col='black', size = 0.1, width = 0.2) + 
  scale_fill_manual(values = c("At.Centre" = color_at_centre, "At.Margin" = color_at_margin)) +
  stat_compare_means(size=3) +
  ylim(c(0,16)) +
  labs(x = "Location", y = "Number of SCNAs")+
  theme(
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),  
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    legend.position = "none"
  )
print (p)

```

# Figure panel (g) -- compare number of SCNAs in less and more necrotic regions

# input data : as above (please run the code block above first!)

```{r}

theme_set(theme_classic())
g <- ggplot(df_modified, aes(Necrotic.Fraction.Interval, Num.Driver.SCNA)) 

# --- print median & sample size ----
print(median(df_modified[which(df_modified$Necrotic.Fraction.Interval == "Less necrotic"), "Num.Driver.SCNA"]))
print(median(df_modified[which(df_modified$Necrotic.Fraction.Interval == "More necrotic"), "Num.Driver.SCNA"]))
print(mean(df_modified[which(df_modified$Necrotic.Fraction.Interval == "Less necrotic"), "Num.Driver.SCNA"]))
print(mean(df_modified[which(df_modified$Necrotic.Fraction.Interval == "More necrotic"), "Num.Driver.SCNA"]))
print(nrow(df_modified[which(df_modified$Necrotic.Fraction.Interval == "Less necrotic"), ]))
print(nrow(df_modified[which(df_modified$Necrotic.Fraction.Interval == "More necrotic"), ]))

# --- VIOLIN ----
p <- g +
  geom_violin(aes(fill= Necrotic.Fraction.Interval), draw_quantiles = c(0.25,0.5,0.75)) +
  #stat_summary(fun.y=mean, fill='black', color='black',geom='point', size=5) +
  #geom_jitter(col='black', size = 0.1) + 
  scale_fill_manual(values = c("Less necrotic" = color_less_necrotic, "More necrotic" = color_more_necrotic)) +
  stat_compare_means(size=3) +
  ylim(c(0,16)) +
  labs(x = "Necrosis of biopsies", y = "Number of SCNAs")+
  theme(
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),  
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    legend.position = "none"
  )
print (p)

```


# Figure panel (h) -- average clone birth times in simulations with or without necrosis

# input data
* simulations with necrosis : ./data_frames/df_necrotic_frac_scna_burden_all_v2.csv
* simulations without necrosis (control) : ./data_frames/df_clone_birth_in_regions_all_CTRL_sims_no_necrosis.csv

```{r}

data_path_control <- "./data_frames/df_clone_birth_in_regions_all_CTRL_sims_no_necrosis.csv"
data_path_necrosis <- "./data_frames/df_clone_birth_in_regions_all.csv"

conds <- c("no necrosis", "necrosis")
paths <- c(data_path_control, data_path_necrosis)

df_compare_clone_birth <- data.frame()

for (k in 1:length(conds)) {
  cond <- conds[k]
  path <- paths[k]
  
  df_clone_birth_in_regions_all <- read.csv(path)
  
  df_clone_birth_in_regions_all$Condition = cond
  
  df_compare_clone_birth <- rbind(df_compare_clone_birth, df_clone_birth_in_regions_all)
}

# --- print median & number of regions ----
df_compare_clone_birth_omit_na <- df_compare_clone_birth %>% drop_na()

print(median(df_compare_clone_birth_omit_na[which(df_compare_clone_birth_omit_na$Condition == "necrosis" & 
          df_compare_clone_birth_omit_na$Location == "At.Centre") , 
   "MeanCloneBirth"]))
print(median(df_compare_clone_birth_omit_na[which(df_compare_clone_birth_omit_na$Condition == "no necrosis" & 
          df_compare_clone_birth_omit_na$Location == "At.Centre") , 
   "MeanCloneBirth"]))
print(mean(df_compare_clone_birth_omit_na[which(df_compare_clone_birth_omit_na$Condition == "necrosis" & 
          df_compare_clone_birth_omit_na$Location == "At.Centre") , 
   "MeanCloneBirth"]))
print(mean(df_compare_clone_birth_omit_na[which(df_compare_clone_birth_omit_na$Condition == "no necrosis" & 
          df_compare_clone_birth_omit_na$Location == "At.Centre") , 
   "MeanCloneBirth"]))

print(nrow(df_compare_clone_birth[
  which(df_compare_clone_birth$Condition == "no necrosis" & 
          df_compare_clone_birth$Location == "At.Centre") , 
  ])
)
print(nrow(df_compare_clone_birth[
  which(df_compare_clone_birth$Condition == "necrosis" & 
          df_compare_clone_birth$Location == "At.Centre") , ])
)
print(nrow(df_compare_clone_birth[
  which(df_compare_clone_birth$Condition == "no necrosis" & 
          df_compare_clone_birth$Location == "At.Margin") , ])
)
print(nrow(df_compare_clone_birth[
  which(df_compare_clone_birth$Condition == "necrosis" & 
          df_compare_clone_birth$Location == "At.Margin") , ])
)

# === VIOLIN ====
theme_set(theme_classic())
g <- ggplot(df_compare_clone_birth_omit_na, aes(x=Location, y=MeanCloneBirth, fill=Condition))

p <- g +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  #stat_summary(fun.y=mean, fill='black', color='black',geom='point', size=5) +
  scale_fill_manual(values = c("no necrosis" = color_less_necrotic, "necrosis" = color_more_necrotic)) +
  #geom_jitter(col='black', size = 0.1) + 
  stat_compare_means(size=3, method = "wilcox") +
  ylim(c(0,1)) +
  labs(x = "Location", y = "Average clone birth time in region")+
  theme(
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),  
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8)
  )
print (p)

```



